Discrete Quantum Theories and Computing

### 2019-02-20 dissertation.tex
dissertation.lyx Secs. 4.1, 4.2, and 4.3 Reviewed...

\begin{figure}[!b]
\noindent \begin{centering}
\includegraphics[viewport=200bp 0bp 1800bp 600bp,width=1.1\columnwidth]{picC}
\par\end{centering}
\caption{\label{bloch.fig}Schematically normalized plots of the elements of
the discrete Bloch sphere, the irreducible single-qubit (two-dimensional)
state vectors with unit norm over the field $\Fpp$. We show the results
for $p=3$, $7$, and $11$. For example, in $\Fppx{3}$, there are
24 vectors of unit norm, but only the 6 inequivalent classes appear
in the plot. The $p+1=4$ equivalent vectors in each class differ
only by a complex discrete phase.}
\end{figure}
In the one-qubit state with coefficients in $\Fpp$, the discrete
analog of the Bloch sphere is constructed by exact analogy to the
continuous case: we first require that the coefficients of the single
qubit basis obey 
\begin{equation}
\fnorm{\ket{\psi_{1}}}=\fnorm{\alpha_{0}}+\fnorm{\alpha_{1}}=1\label{DQA2.eq}
\end{equation}
in the discrete field. We show that there are $p\left(p^{2}-1\right)$
such values later in the general theorem, Proposition~\ref{appU.sec}.
Given this requirement, which is similar in form to the conservation
of probability, but not as useful due to the lack of orderable probability
values, we can immediately conclude that the discrete analog of the
Hopf fibration is again 
\begin{eqnarray}
X & = & 2\,\mathrm{Re}\ \alpha_{0}\alpha_{1}^{*}=2x_{0}x_{1}+2y_{0}y_{1}\,,\nonumber \\
Y & = & 2\,\mathrm{Im}\ \alpha_{0}\alpha_{1}^{*}=2x_{1}y_{0}-2x_{0}y_{1}\,,\label{DQA6.eq}\\
Z & = & \fnorm{\alpha_{0}}-\fnorm{\alpha_{1}}={x_{0}}^{2}+{y_{0}}^{2}-{x_{1}}^{2}-{y_{1}}^{2}\,.\nonumber 
\end{eqnarray}
but now with all computations in\hspace{0bp}$\pmod{p}$. At this
point, one can simply write down all possible discrete values for
the complex numbers $\left(\alpha_{0},\alpha_{1}\right)$ satisfying
Eq.~(\ref{DQA2.eq}) and enumerate those that project to the same
value of $\left(X,Y,Z\right)$. This equivalence class is the discrete
analog of the circle in the complex plane that was eliminated in the
continuous case. In Proposition~\ref{appH.sec}, we show that $p+1$
discrete values of $\left(\alpha_{0},\alpha_{1}\right)$ with unit
norm map to the same point under the Hopf map \Eq{DQA6.eq}; %
\begin{comment}
\todo{Does Proposition~\ref{appH.sec} really show this? }
\end{comment}
we may think of these as discrete circles or projective lines of equivalent,
physically indistinguishable, complex phase. The surviving $p\left(p-1\right)$
values of $\left(\alpha_{0},\alpha_{1}\right)$ correspond to irreducible
physical states of the discrete single qubit system. Thus, for example,
choosing the underlying field to be $\Fppx{3}$, there are exactly
6 single-qubit state vectors to populate the Bloch sphere; the four
equivalent phase-multiples mapping to each of the six points on the
$\Fppx{3}$ Bloch sphere are collapsed and regarded as physically
indistinguishable. In Figure~\ref{bloch.fig}, we plot the irreducible
states on the Bloch sphere for $p=3$, $7$, and $11$. Note that
the Cartesian lengths of the real vectors corresponding to the points
on the Bloch sphere vary considerably due to the nature of discrete
fields; we have artificially normalized them to a ``continuous world''
unit radius sphere for conceptual clarity.

>>>>

In the one-qubit state with coefficients in $\Fpp$, the discrete
analog of the Bloch sphere is constructed by exact analogy to the
continuous case: we first require that the coefficients of the single
qubit basis obey 
\begin{equation}
\fnorm{\ket{\psi_{1}}}=\fnorm{\alpha_{0}}+\fnorm{\alpha_{1}}=1\label{DQA2.eq}
\end{equation}
in the discrete field. We show that there are $p\left(p^{2}-1\right)$
such values later in the general theorem, Proposition~\ref{appU.sec}.
Given this requirement, which is similar in form to the conservation
of probability, but not as useful due to the lack of orderable probability
values, we can immediately conclude that the discrete analog of the
Hopf fibration is again 
\begin{eqnarray}
X & = & 2\,\mathrm{Re}\ \alpha_{0}\alpha_{1}^{*}=2x_{0}x_{1}+2y_{0}y_{1}\,,\nonumber \\
Y & = & 2\,\mathrm{Im}\ \alpha_{0}\alpha_{1}^{*}=2x_{1}y_{0}-2x_{0}y_{1}\,,\label{DQA6.eq}\\
Z & = & \fnorm{\alpha_{0}}-\fnorm{\alpha_{1}}={x_{0}}^{2}+{y_{0}}^{2}-{x_{1}}^{2}-{y_{1}}^{2}\,,\nonumber 
\end{eqnarray}
but now with all computations in\hspace{0bp}$\pmod{p}$. At this
point, one can simply write down all possible discrete values for
the complex numbers $\left(\alpha_{0},\alpha_{1}\right)$ satisfying
Eq.~(\ref{DQA2.eq}) and enumerate those that project to the same
value of $\left(X,Y,Z\right)$. This equivalence class is the discrete
analog of the circle in the complex plane that was eliminated in the
continuous case. In Proposition~\ref{appH.sec}, we show that $p+1$
discrete values of $\left(\alpha_{0},\alpha_{1}\right)$ with unit
norm map to the same point under the Hopf map \Eq{DQA6.eq}; %
\begin{comment}
\todo{Does Proposition~\ref{appH.sec} really show this? }
\end{comment}
we may think of these as discrete circles or projective lines of equivalent,
physically indistinguishable, complex phase. The surviving $p\left(p-1\right)$
values of $\left(\alpha_{0},\alpha_{1}\right)$ correspond to irreducible
physical states of the discrete single qubit system. Thus, for example,
choosing the underlying field to be $\Fppx{3}$, there are exactly
6 single-qubit state vectors to populate the Bloch sphere; the four
equivalent phase-multiples mapping to each of the six points on the
$\Fppx{3}$ Bloch sphere are collapsed and regarded as physically
indistinguishable. In Figure~\ref{bloch.fig}, we plot the irreducible
states on the Bloch sphere for $p=3$, $7$, and $11$. Note that
the Cartesian lengths of the real vectors corresponding to the points
on the Bloch sphere vary considerably due to the nature of discrete
fields; we have artificially normalized them to a ``continuous world''
unit radius sphere for conceptual clarity.
\begin{figure}[!b]
\noindent \begin{centering}
\includegraphics[viewport=200bp 0bp 1800bp 600bp,width=1.1\columnwidth]{picC}
\par\end{centering}
\caption{\label{bloch.fig}Schematically normalized plots of the elements of
the discrete Bloch sphere, the irreducible single-qubit (two-dimensional)
state vectors with unit norm over the field $\Fpp$. We show the results
for $p=3$, $7$, and $11$. For example, in $\Fppx{3}$, there are
24 vectors of unit norm, but only the 6 inequivalent classes appear
in the plot. The $p+1=4$ equivalent vectors in each class differ
only by a complex discrete phase.}
\end{figure}
